---
title: "<center>Installing and Configuring R and Rstudio with ASReml-R</center>"
author: 
- <center>Biometry Hub, University of Adelaide<center>
- <center>biometrytraining@adelaide.edu.au<center>
date: "<center>Last updated: `r format(Sys.Date(), format='%B %Y')`</center>"
output:
  html_document:
    toc: yes
    toc_float: yes
    keep_md: true
    highlight: textmate
  pdf_document:
    fig_caption: yes
    toc: yes
linkcolor: blue
urlcolor: blue
params:
    rlatest: "4.1.2"
---

```{r echo=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```


---

# Introduction  

These instructions assume you already have a working version of R on your computer. If this is not yet the case, please follow these instructions: https://biometryhub.github.io/installing-r/installing-r.html

This guide uses Rstudio, but that is optional - it will work equally well with any alternative R development environment.

This guide assumes you are running a relatively recent version of Microsoft Windows (e.g. Windows 8/8.1/10) or Mac OS X/macOS (10.6+). The guide will likely work for Debian based Linux distributions as well (E.g. Ubuntu flavours), however it is not extensively tested under those scenarios.

# ASReml-R Installation

## biometryassist

You will need the `biometryassist` package, which can be installed from CRAN via

```{r}
install.packages("biometryassist")
```



## ASReml-R

Once `biometryassist` is installed, the latest version of ASReml-R can then be installed with the commands

```{r}
library(biometryassist)
install_asreml()
```

The `install_asreml()` function comes from the `biometryassist` package, hence the need to install and load that first.

All going well, you should get a confirmation that ASReml-R is installed successfully. 

# Done!  
Congratulations! If you have made it this far, you are now set up and ready for the workshop.  


---


# Further resources {#Further}
The above will get your basic setup ready but here are some links if you are interested in reading a bit further.

- R project homepage: <https://www.r-project.org/>
- RStudio homepage: <https://www.rstudio.com/products/rstudio/>
- How to Use RStudio: <https://www.rstudio.com/ide/docs/>
- R Installation and Administration: <https://cran.r-project.org/doc/manuals/R-admin.html>
- R FAQ: <https://cran.r-project.org/doc/FAQ/R-FAQ.html>
- More about add-on packages in the R Installation and Administration Manual:  <https://cran.r-project.org/doc/manuals/R-admin.html#Add_002don-packages>

